---
title: "SnpEff_script"
author: "Dani"
date: "6 de junio de 2017"
output: html_document
---

#Search for the Lynx pardinus database
Step 1. Search for the Lynx pardinus assembly database in the program's pre-built database. As of June the 6th, 2017, the Lynx pardinus genome isn't included in the snpEff database. Once the snpEff authors include it, a newer version of the database should be downloaded (ask the tech group). A second option would be building our own Lynx pardinus database, although it seems complicated.


```{r Search for the Lynx pardinus database, eval=FALSE, engine='bash'}

S_PATH=/opt/snpEff #software path
I_PATH= #write here the input path
O_PATH=/home/dkleinman/datos/snpEff #output path

java -jar /opt/snpEff/snpEff.jar databases | grep -i pardinus

```


#Build the Lynx pardinus genome database
Step 2. 

```{r Build the Lynx pardinus genome database, eval=FALSE, engine='bash'}

/home/dkleinman/
vi snpEffect.config 
#Originally the config was just in the software folder and I didn't have writing permission. If this is the only config file available, writing permission is required, and when annotating later on, the file should be called using the -c command followed by the path to the file.
#However, in my case I believe the tech group created a copy of the file in my folder, and this is the one that I was able to edit. I added the following two lines:

# Lynx_pardinus
LYPA.23.genome : Iberian lynx 

```


#Tutorial annotation
Step 3. Annotate one of the examples that comes with the software

```{r Search for the Lynx pardinus database, eval=FALSE, engine='bash'}

java -Xmx16g -jar $S_PATH/snpEff.jar GRCh37.75 -s $O_PATH/toys/test.chr22.ann $S_PATH/examples/test.chr22.vcf > $O_PATH/toys/test.chr22.ann.vcf

```

