---
title: "SnpEff_script"
author: "Dani"
date: "6 de junio de 2017"
output: html_document
---

http://snpeff.sourceforge.net/SnpEff_manual.html #I'll be following this manual for all configuration purposes

#0a: Define paths.

```{r Define paths, eval=FALSE, engine='bash'}

S_PATH=/opt/snpEff #software path
C_PATH=/home/dkleinman/datos/snpEff #config file path
O_PATH=/home/dkleinman/datos/snpEff #output path
I_PATH=/home/GRUPOS/grupolince/immunocapture/prueba_highdiv #immunocapture path
V_PATH=/home/GRUPOS/grupolince/lynx_genomes_5x/VCFs_Dani #VCFs path

```


#0b: Search for the Lynx pardinus database.
Search for the Lynx pardinus assembly database in the program's pre-built database. As of June the 6th, 2017, the Lynx pardinus genome isn't included in the snpEff database. A second option would be building our own Lynx pardinus database.

```{r Search for the Lynx pardinus database, eval=FALSE, engine='bash'}

java -jar /opt/snpEff/snpEff.jar databases | grep -i pardinus

```


#0c: Build the Lynx pardinus genome database.
In the end we opt to build our own database since we don't know when they will get themselves to add it. This step should be omitted if the desired database was found in the previous step.

##Add entry to the config file

```{r Build the Lynx pardinus genome database, eval=FALSE, engine='bash'}

#Originally the config was just in the software folder and I didn't have writing permission. If this is the only config file available, writing permission is required, and when annotating later on, the file should be called using the -c command followed by the path to the file.
#However, in my case I believe the tech group created a copy of the file in my folder after I sent them an e-mail, and this is the one that I was able to edit.

cd /home/dkleinman/
mv snpEff.config $C_PATH #I move the config file that appeared in my folder to a subfolder that I created for snpEff
vi snpEff.config  #initiate the editing process

#Following the manual, I added the following two lines:

# Lynx_pardinus
LYPA.23.genome : Iberian lynx #from now on, LYPA.23 is the code for the Lynx pardinus reference genome (in snpEff)

# Lynx_pardinus, detailed annotation
LYPA.23b.genome : Iberian lynx #LYPA.23b is the code for the highly detailed annotation of the Lynx pardinus reference genome (in snpEff)

```

##Create directory and move files

```{r Build the Lynx pardinus genome database, eval=FALSE, engine='bash'}

#First for the regular annotation:
mkdir $S_PATH/data/LYPA.23 #create a directory inside the software's dependencies whose name matches the code
cd $S_PATH/data/LYPA.23

scp /GRUPOS/grupolince/Lyp_annotation_Apr14_final/LYPA23C.all.fix.nr.gff3 $S_PATH/data/LYPA.23/ #copy the annotation file (can be gff or gtf) to the newly created directory. This gff file includes CDS, introns, exons and genes, so it's very basic. A more complex version that Maria created which includes lncRNAs, etc., could be tested in the future.
mv LYPA23C.all.fix.nr.gff3 genes.gff #rename the file as the tutorial indicates

mkdir $S_PATH/data/genomes #create a directory inside the software's dependencies called genomes
cd $S_PATH/data/genomes
scp /home/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/lp23.fa $S_PATH/data/genomes #copy the reference genome fasta to the new genomes directory
mv lp23.fa LYPA.23.fa #rename the file so that it matches the code




#Second, for the detailed annotation:
mkdir $C_PATH/data/LYPA.23b #create a directory inside the software's dependencies whose name matches the code
cd $C_PATH/data/LYPA.23b

scp /GRUPOS/grupolince/Lyp_annotation_Apr14_final/LYPA23C.GENE.mRNA.CDS.GENE_promoters.GENE_introns.UTRs.ncRNA.lncRNA.lncRNA_introns.lncRNA_promoters.intergenic.nr.gff3 $C_PATH/data/LYPA.23b/ #copy the annotation file (can be gff or gtf) to the newly created directory. This gff file is very detailed and includes CDS, introns, exons, genes, and many more.
mv LYPA23C.GENE.mRNA.CDS.GENE_promoters.GENE_introns.UTRs.ncRNA.lncRNA.lncRNA_introns.lncRNA_promoters.intergenic.nr.gff3 genes.gff #rename the file as the tutorial indicates

cd $C_PATH/data/genomes
scp LYPA.23.fa LYPA.23b.fa #copy the reference genome and rename it so that it also matches the detailed annotation entry

```

##Build the database

```{r Build the Lynx pardinus genome database, eval=FALSE, engine='bash'}

cd $C_PATH
screen -S build_snpEff_db_detailed #open a dettachable screen in case the database building takes too long
script build_snpEff_db_detailed.txt #initiate the log file

S_PATH=/opt/snpEff #redefine the variable, since we're inside a script
C_PATH=/home/dkleinman/datos/snpEff #redefine the variable, since we're inside a script

cd $S_PATH
java -jar snpEff.jar build -gff3 -v LYPA.23b -c $C_PATH/snpEff.config -dataDir $C_PATH/data #build the database. Use the -gff3 command for gff files and -gtf22 for gtf files. Use -v for verbose (expanded information on the processes and the warnings/errors that may appear). Use -c to indicate the path to my own config file. Then use -dataDir to override the data directory from the config file (by default the software thinks that the data folder with the genome and the genes files is located where config is, so it's necessary to give it the correct path). Another option would be running it from the config folder and indicating instead the 

ctrl + D #terminate the script
ctrl + D #terminate the screen

scp -r $S_PATH/data $C_PATH #afterwards I realize anyone can access the data folder so I copy it to my own folder and then I remove the stuff I created inside the original data folder
cd $S_PATH/data
rm -r LYPA.23/
rm -r genomes/

```

#1: Tutorial annotation.
Annotate one of the examples that comes with the software

```{r Tutorial annotation, eval=FALSE, engine='bash'}

java -Xmx16g -jar $S_PATH/snpEff.jar GRCh37.75 -s $O_PATH/toys/test.chr22.ann $S_PATH/examples/test.chr22.vcf > $O_PATH/toys/test.chr22.ann.vcf

```

#2: Toy annotation using the immunocapture VCF.
Annotate the immunocapture VCF 

```{r Toy annotation using the immunocapture VCF, eval=FALSE, engine='bash'}

cd $C_PATH
screen -S immuno_annot_test1 #open a dettachable screen in case the test takes too long

script immunocapture/immuno_annot_test1.txt #initiate the log file

S_PATH=/opt/snpEff #redefine the variable, since we're inside a script
C_PATH=/home/dkleinman/datos/snpEff #redefine the variable, since we're inside a script
O_PATH=/home/dkleinman/datos/snpEff #redefine the variable, since we're inside a script
I_PATH=/home/GRUPOS/grupolince/immunocapture/prueba_highdiv #write here the input path. This folder only exists in server A, so keep that in mind (I was looking for it for the longest time in server B >.<)

java -Xmx16g -jar $S_PATH/snpEff.jar LYPA.23 -v -s $O_PATH/immunocapture/immunocapture_test1.ann.html $I_PATH/x_lx_xx_n201_filtered.vcf > $O_PATH/immunocapture/immunocapture_test1.ann.vcf #run this code from the directory where the config is located, and from server A (I_PATH doesn't exist in server B). I used the filtered version of the vcf which includes all lynx individuals in the immunocapture; there's also a raw version which can be tested.

ctrl + D #terminate the script
ctrl + D #terminate the screen

scp dkleinman@genomics-b.ebd.csic.es:/home/dkleinman/datos/snpEff/immunocapture/immunocapture_test1.ann* /Users/Dani/ownCloud/backup/annotation/snpEff_results/immunocapture/ #execute this command out of the server to import to the pc the output files of this toy run

```

#3a: Annotation (basic) using the contemporary lp and ll VCFs.
Annotate the contemporary VCFs.

```{r Annotation using the contemporary lp and ll VCFs, eval=FALSE, engine='bash'}

cd $C_PATH
screen -S c_ll_lp_perpop_standard_filter_annotation.log #open a dettachable screen in case the test takes too long
script c_ll_lp_perpop_standard_filter_annotation.log #initiate the log file

S_PATH=/opt/snpEff #software path
C_PATH=/home/dkleinman/datos/snpEff #config file path
O_PATH=/home/dkleinman/datos/snpEff #output path
I_PATH=/home/GRUPOS/grupolince/immunocapture/prueba_highdiv #immunocapture path
V_PATH=/home/GRUPOS/grupolince/lynx_genomes_5x/VCFs_Dani #VCFs path

cd $V_PATH
shopt -s extglob #the extglob shell option gives you more powerful pattern matching in the command line.
POPLIST=($(ls l*[^pv]_perpop_standard_filter.vcf)) #selects all standard-filtered populational VCFs except for País Vasco
for pop in "${POPLIST[@]}"
  do
  echo "${pop}"
  #done
  cd $C_PATH
  java -Xmx16g -jar $S_PATH/snpEff.jar LYPA.23 -v -csvStats $O_PATH/c_ll_lp_perpop_standard_filter_annotation/${pop/.vcf/.ann.csv} $V_PATH/${pop} > $O_PATH/c_ll_lp_perpop_standard_filter_annotation/${pop/.vcf/.ann.vcf} #run this code from the directory where the config is located.
done
shopt -u extglob #disable extglob

scp dkleinman@genomics-b.ebd.csic.es:/home/dkleinman/datos/snpEff/c_ll_lp_perpop_standard_filter_annotation/l*csv /Users/Dani/ownCloud/backup/annotation/snpEff_results/c_ll_lp_perpop_standard_filter_annotation/
scp dkleinman@genomics-b.ebd.csic.es:/home/dkleinman/datos/snpEff/c_ll_lp_perpop_standard_filter_annotation/*html /Users/Dani/ownCloud/backup/annotation/snpEff_results/c_ll_lp_perpop_standard_filter_annotation/
#scp dkleinman@genomics-b.ebd.csic.es:/home/dkleinman/datos/snpEff/c_ll_lp_perpop_standard_filter_annotation/l*txt /Users/Dani/ownCloud/backup/annotation/snpEff_results/c_ll_lp_perpop_standard_filter_annotation/

```

#3b: Annotation (detailed) using the contemporary lp and ll VCFs.
Annotate the contemporary VCFs.

```{r Annotation using the contemporary lp and ll VCFs, eval=FALSE, engine='bash'}

screen -S c_ll_lp_perpop_standard_filter_detailed_annotation.log #open a dettachable screen in case the test takes too long
script /home/dkleinman/datos/snpEff/c_ll_lp_perpop_standard_filter_detailed_annotation/c_ll_lp_perpop_standard_filter_detailed_annotation.log #initiate the log file

S_PATH=/opt/snpEff #software path
C_PATH=/home/dkleinman/datos/snpEff #config file path
O_PATH=/home/dkleinman/datos/snpEff #output path
I_PATH=/home/GRUPOS/grupolince/immunocapture/prueba_highdiv #immunocapture path
V_PATH=/home/GRUPOS/grupolince/lynx_genomes_5x/VCFs_Dani #VCFs path

cd $V_PATH
shopt -s extglob #the extglob shell option gives you more powerful pattern matching in the command line.
POPLIST=($(ls l*[^pv]_perpop_standard_filter.vcf)) #selects all standard-filtered populational VCFs except for País Vasco
for pop in "${POPLIST[@]}"
  do
  echo "${pop}"
  #done
  cd $C_PATH
  java -Xmx16g -jar $S_PATH/snpEff.jar LYPA.23b -v -csvStats $O_PATH/c_ll_lp_perpop_standard_filter_detailed_annotation/${pop/.vcf/.ann.csv} $V_PATH/${pop} > $O_PATH/c_ll_lp_perpop_standard_filter_detailed_annotation/${pop/.vcf/.ann.vcf} #run this code from the directory where the config is located.
done
shopt -u extglob #disable extglob


```





