---
title: "substitutions_lynx_genus_VCFs"
author: "Dani"
date: "20 de febrero de 2018"
output: html_document
---


#0a: Define paths.

```{r Define paths, eval=FALSE, engine='bash'}

B_PATH=/home/GRUPOS/grupolince/lynx_genomes_5x/BAM_files_final #BAMs path
G_PATH=/GRUPOS/grupolince/lynx_genomes_5x/gVCFs #gVCFs path
V_PATH=/home/dkleinman/datos/VCFs_Dani #VCFs path
REF=/home/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/lp23.fa #path to reference genome
GATK=/opt/GATK-3.7/GenomeAnalysisTK.jar #GATK software path
BCF=/opt/bcftools-1.6/bcftools #BCFtools software path

```

#0b: Prepare reference genome. Prepare two files with data from the reference genome: a dictionary with contig names and sizes, and a fasta index file. This step should only be performed once (per reference genome).

```{r Prepare reference genome, eval=FALSE, engine='bash'}

#Karolina's code to prepare the reference genome. This should only be performed once per reference genome.
cd /home/GRUPOS/grupolince/reference_genomes/lynx_pardinus_genome/
java -jar /home/tmp/Software/Picard/picard-tools-1.66/CreateSequenceDictionary.jar \
 R=lp23.fa \
 O=lp23.dict
samtools faidx lp23.fa

```

#1a: Produce gVCF files. Perform a calling per sample to produce a gVCF file with variant information for every position in the genome (variant or not).

```{r Produce gVCF files, eval=FALSE, engine='bash'}

#We'll be using one 25x sample from each species: c_lp_sm_0298_recal_round-1_25x, c_ll_ki_0090_recal_round-1_25x, c_lc_zz_0001_recal_round-1_25x, c_lr_zz_0001_recal_round-1_25x. The gVCF for lp_sm_0298 was already created months ago as part of the following loop (in the c_lp_ll_VCF script):

cd $B_PATH/genome_project_samples_25x #this works for the 25x samples
POP=("*_lp")
for pop in ${POP[@]}
  do
  echo "${pop}"
  ls ${pop}_*_recal_round-1_25x.bam > ${pop/*_lp/lp}_recal_round-1_25x.bam.list;
  INPUT_BAMS_FOR_CALLING=($(cat ${pop/*_lp/lp}_recal_round-1_25x.bam.list)) 
  for id in ${INPUT_BAMS_FOR_CALLING[@]}
    do
    echo "${id}"
    java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -jar $GATK \
    -T HaplotypeCaller \
    -R $REF \
    -I ${id} \
    --emitRefConfidence GVCF \
    -o $G_PATH/genome_project_samples_25x/${id/.bam/.g.vcf.gz}
    done
  done


#The remaining three gVCFs were produced as follows:
    
#This was only used to produce the gVCF file for the original (not subsampled) Macrogen Kirov sample:
cd $B_PATH/MACROGEN_samples_25x
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -Xms16g -Xmx32g -jar $GATK \
    -T HaplotypeCaller \
    -R $REF \
    -I c_ll_ki_0090_recal_round-1_25x.bam \
    --emitRefConfidence GVCF \
    -o $G_PATH/macrogen_samples_25x/c_ll_ki_0090_recal_round-1_25x_bis.g.vcf.gz

#This was only used to produce the gVCF file for the original (not subsampled) Macrogen Lynx canadensis sample:
cd $B_PATH/MACROGEN_samples_25x
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -Xms16g -Xmx32g -jar $GATK \
    -T HaplotypeCaller \
    -R $REF \
    -I c_lc_zz_0001_recal_round-1_25x.bam \
    --emitRefConfidence GVCF \
    -o $G_PATH/macrogen_samples_25x/c_lc_zz_0001_recal_round-1_25x.g.vcf.gz
    
#This was only used to produce the gVCF file for the original (not subsampled) Macrogen Lynx rufus sample:
cd $B_PATH/MACROGEN_samples_25x
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -Xms16g -Xmx32g -jar $GATK \
    -T HaplotypeCaller \
    -R $REF \
    -I c_lr_zz_0001_recal_round-1_25x.bam \
    --emitRefConfidence GVCF \
    -o $G_PATH/macrogen_samples_25x/c_lr_zz_0001_recal_round-1_25x.g.vcf.gz

```

#1b: Perform sanity checks on gVCFs. Perform various sanity checks on all gVCFs.

```{r Perform sanity checks on gVCFs, eval=FALSE, engine='bash'}

cd $G_PATH
shopt -s extglob #the extglob shell option gives you more powerful pattern matching in the command line.
SAMPLELIST=($(ls c_*.g.vcf.gz | cut -d "." -f1 | sort | uniq))
rm c_gVCF_raw.stats
echo "sample_name,total_SNPs,homoz_ref,heteroz,homoz_alt,unaccounted" > c_gVCF_raw.stats 
for sample in "${SAMPLELIST[@]}"
  do
  echo "${sample}"
  #done
  NAME="${sample}"
  TOTAL_SNPS="$(zgrep -v '#' "$G_PATH/${sample}".g.vcf.gz | wc -l)"
  echo $TOTAL_SNPS
  TOTAL_00="$(zgrep -v '#' "$G_PATH/${sample}".g.vcf.gz | zgrep '0/0:' | wc -l)"
  echo $TOTAL_00
  TOTAL_01="$(zgrep -v '#' "$G_PATH/${sample}".g.vcf.gz | zgrep '0/1:' | wc -l)"
  echo $TOTAL_01
  TOTAL_11="$(zgrep -v '#' "$G_PATH/${sample}".g.vcf.gz | zgrep '1/1:' | wc -l)"
  echo $TOTAL_11
  UNACCOUNTED="$(($TOTAL_SNPS - $TOTAL_00 - $TOTAL_01 - $TOTAL_11))"
  echo $UNACCOUNTED
  echo "$NAME,$TOTAL_SNPS,$TOTAL_00,$TOTAL_01,$TOTAL_11,$UNACCOUNTED" >> c_gVCF_raw.stats
  done
shopt -u extglob #disable extglob

#Save locally the .stats file
scp dkleinman@genomics-b.ebd.csic.es:$G_PATH/c_gVCF_raw.stats /Users/Dani/ownCloud/backup/annotation/gVCFs_analysis/

```

#1c: Analyse the sanity checks file. Perform per individual analyses on the sanity checks file.

```{r Analyse the sanity checks file}
library("readr")
library("dplyr")
library("ggplot2")

local_repo <- file.path("/Users/Dani/ownCloud/backup/annotation/gVCFs_analysis/")

sanity_checks <- read_csv("/Users/Dani/ownCloud/backup/annotation/gVCFs_analysis/c_gVCF_raw.stats")
sanity_checks
rowSums(sanity_checks[,-c(1,2)])==sanity_checks[,2] #check if the sum of AA, AB and BB SNPs equal the total SNP count (for each individual)
pop_summary_sc <- sanity_checks %>% 
  mutate(species=substr(sanity_checks$sample_name,3,4),population=substr(sanity_checks$sample_name,6,7)) %>%
  group_by(species,population) %>%
  summarise(N=n(),mean_total_SNPs=mean(total_SNPs),mean_homoz_ref=mean(homoz_ref),mean_heteroz=mean(heteroz),mean_homoz_alt=mean(homoz_alt),mean_unaccounted=mean(unaccounted))
pop_summary_sc

plot_pop_total_SNPs <- ggplot(pop_summary_sc, aes(population,mean_total_SNPs)) + geom_col() + facet_grid(. ~ species,scales="free_x",space="free_x")
plot_pop_total_SNPs

plot_pop_heteroz <- ggplot(pop_summary_sc, aes(population,mean_heteroz)) + geom_col() + facet_grid(. ~ species,scales="free_x",space="free_x")
plot_pop_heteroz

plot_indiv_total_SNPs <- ggplot(sanity_checks, aes(x=substr(sample_name,0,12),y=total_SNPs,fill=substr(sample_name,0,7))) + geom_col() + theme(axis.text.x=element_text(angle=90,hjust=0,size=6,colour="black")) 
plot_indiv_total_SNPs

plot_indiv_heteroz <- ggplot(sanity_checks[grep("c_ll",sanity_checks$sample_name),], aes(x=substr(sample_name,0,12),y=heteroz,fill=substr(sample_name,6,7))) + 
  geom_col() +
  ggtitle("0/1 positions per sample (in gVCF)") +
  theme_bw() +
  labs(x="sample",y="N",fill="population") +
  theme(plot.title=element_text(hjust=0.5),
        axis.text.x=element_text(angle=90, hjust=0, size=6,colour="black"),
        legend.key.size=unit(0.4,"cm")
  )
plot_indiv_heteroz
ggsave("gVCFs_SNPs_per_sample.pdf", width=30, height=10, units="cm", device="pdf", path=local_repo)

```

#2a: Combine gVCF into VCF. Run joint genotyping of all desired gVCF files to produce a multisample VCF file.

```{r Combine gVCF into VCF, eval=FALSE, engine='bash'}

#Create an array with the path to all desired files:
GVCF_LIST=($G_PATH/genome_project_samples_25x/c_lp_sm_0298_recal_round-1_25x.g.vcf.gz $G_PATH/macrogen_samples_25x/c_ll_ki_0090_recal_round-1_25x.g.vcf.gz $G_PATH/macrogen_samples_25x/c_lc_zz_0001_recal_round-1_25x.g.vcf.gz $G_PATH/macrogen_samples_25x/c_lr_zz_0001_recal_round-1_25x.g.vcf.gz)

#Combine gVCF files into a VCF that conserves all sites, including non-variable sites (this will be important later on to compare callable sites and uncovered sites with total sites):
cd $G_PATH
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -jar $GATK \
-T GenotypeGVCFs \
-R $REF \
$(for var in ${GVCF_LIST[@]}; do echo -V ${var}" ";done) \
--includeNonVariantSites \
-o $V_PATH/c_lp_ll_lc_lr_all_sites.vcf

#Combine gVCF files into a VCF with all variant sites:
cd $G_PATH
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -jar $GATK \
-T GenotypeGVCFs \
-R $REF \
$(for var in ${GVCF_LIST[@]}; do echo -V ${var}" ";done) \
-o $V_PATH/c_lp_ll_lc_lr_variant_sites.vcf

#Get the number of SNPs of the VCFs:
grep -v '#' c_lp_ll_lc_lr_all_sites.vcf | wc -l #2412763898
grep -v '#' c_lp_ll_lc_lr_variant_sites.vcf | wc -l #30079077

```

#2b: Fix sample names in the VCF.

```{r Fix sample names in the VCF, eval=FALSE, engine='bash'}

#It's important to rename those samples in the global VCF that don't have the proper name (the 3 that were sequenced at Macrogen).

#First, do it for the variants-only VCF:
cd $V_PATH
$BCF query -l c_lp_ll_lc_lr_variant_sites.vcf #checks sample names
#!/bin/bash
#cat << "EOF" > list_to_remove.txt #Unmark this line when running it (I marked it in R because it doesn't understand it and messes the colours of subsequent lines). It should include all incorrect names.
LC1_rgsm c_lc_zz_0001
LL90_rgsm c_ll_ki_0090
LR1_rgsm c_lr_zz_0001
EOF

$BCF reheader -s list_to_remove.txt -o c_lp_ll_lc_lr_variant_sites_renamed.vcf c_lp_ll_lc_lr_variant_sites.vcf #copy the VCF and rename the wrong named samples in the new VCF

cd $V_PATH
$BCF query -l c_lp_ll_lc_lr_variant_sites_renamed.vcf #checks if samples have been renamed properly


#Then, repeat it for the all-sites VCF:
cd $V_PATH
$BCF query -l c_lp_ll_lc_lr_all_sites.vcf #checks sample names
#!/bin/bash
#cat << "EOF" > list_to_remove.txt #Unmark this line when running it (I marked it in R because it doesn't understand it and messes the colours of subsequent lines). It should include all incorrect names.
LC1_rgsm c_lc_zz_0001
LL90_rgsm c_ll_ki_0090
LR1_rgsm c_lr_zz_0001
EOF

$BCF reheader -s list_to_remove.txt -o c_lp_ll_lc_lr_all_sites_renamed.vcf c_lp_ll_lc_lr_all_sites.vcf #copy the VCF and rename the wrong named samples in the new VCF

cd $V_PATH
$BCF query -l c_lp_ll_lc_lr_all_sites_renamed.vcf #checks if samples have been renamed properly

```

#3: Split the VCF into per species VCFs. Generate a VCF for each individual.

```{r Split the VCF into per species VCFs, eval=FALSE, engine='bash'}

#Split the VCF in order to get one for each species. Contrary to our objective with the lp & ll VCF, here it's important to keep all substitutions.

cd $G_PATH
GVCF_LIST=($G_PATH/genome_project_samples_25x/c_lp_sm_0298_recal_round-1_25x.g.vcf.gz $G_PATH/macrogen_samples_25x/c_ll_ki_0090_recal_round-1_25x.g.vcf.gz $G_PATH/macrogen_samples_25x/c_lc_zz_0001_recal_round-1_25x.g.vcf.gz $G_PATH/macrogen_samples_25x/c_lr_zz_0001_recal_round-1_25x.g.vcf.gz)
for i in ${GVCF_LIST[@]}
  do
  echo "${i}"
  INPUT=$(echo "${i}" | cut -d$"/" -f8 | cut -d$"_" -f1,2,3,4)
  echo $INPUT
  $BCF view -s $INPUT -Ov -o $V_PATH/"$INPUT"_recal_round-1_25x.vcf $V_PATH/c_lp_ll_lc_lr_all_sites_renamed.vcf
  done

grep -v '#' c_lp_sm_0298_recal_round-1_25x.vcf | wc -l #2412763898
grep -v '#' c_ll_ki_0090_recal_round-1_25x.vcf | wc -l #2412763898
grep -v '#' c_lc_zz_0001_recal_round-1_25x.vcf | wc -l #2412763898
grep -v '#' c_lr_zz_0001_recal_round-1_25x.vcf | wc -l #2412763898

#Check whether the reference allele has changed between VCFs. To this end, let's compare all SNPs common to both species' variants list.
grep -v '#' $V_PATH/c_lr_zz_0001_recal_round-1_25x.vcf | cut -d$'\t' -f1,2,4,5 | awk '{print $1"_"$2" "$3" "$4}' | sort -k 1,1 > $V_PATH/c_lr_zz_0001_recal_round-1_25x_variants.txt #creates variants list for lr.
grep -v '#' $V_PATH/c_ll_ki_0090_recal_round-1_25x.vcf | cut -d$'\t' -f1,2,4,5 | awk '{print $1"_"$2" "$3" "$4}' | sort -k 1,1 > $V_PATH/c_ll_ki_0090_recal_round-1_25x_variants.txt #creates variants list for ll.
join -1 1 -2 1 -e0 -o'0,1.2,1.3,2.2,2.3' $V_PATH/c_lr_zz_0001_recal_round-1_25x.vcf $V_PATH/c_ll_ki_0090_recal_round-1_25x_variants.txt > c_lr_zz_0001_c_ll_ki_0090_joined_variants.txt #joins both lists and outputs only the common SNPs.
awk '{if($2 == $4) print 1; else print 0; }' c_lr_zz_0001_c_ll_ki_0090_joined_variants.txt | sort | uniq #this should only return the value "1" when all SNPs share the same reference allele, or both "0" and "1" if some SNPs have different reference alleles between species. In this case it only returns "1", so everything is fine.
cat c_lr_zz_0001_c_ll_ki_0090_joined_variants.txt | awk '{if($2 == $4) print $0" 1"; else print $0" 0"; }' c_lr_zz_0001_c_ll_ki_0090_joined_variants.txt | awk '$6 == 0' #in case there's any row where the reference allele has changed, return those rows. In this case, none are returned.

```

#4: Get substitutions for any pair of species.
##Lynx lynx - Lynx rufus.
###Obtain joint VCF for eurasian and rufus. Extract data for the two species in order to catalogue substitutions.

```{r Obtain joint VCF for eurasian and rufus, eval=FALSE, engine='bash'}

#Obtain VCF for the eurasian and the rufus:
cd $V_PATH
$BCF view -s c_ll_ki_0090,c_lr_zz_0001 -Ov -o $V_PATH/c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x.vcf $V_PATH/c_lp_ll_lc_lr_all_sites_renamed.vcf

grep -v '#' c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x.vcf | wc -l #2412763898

```

###Extract SNPs from the VCF. Subset the VCF files in order to keep only SNP variants.

```{r Extract SNPs from the VCF, eval=FALSE, engine='bash'}

#During this step, all INDELs will be dropped from the VCF (not entirely true... INDELs where one variant is monoallelic will remain).
cd $V_PATH
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -Xms16g -Xmx32g -jar $GATK \
  -T SelectVariants \
  --selectTypeToExclude INDEL \
  -R $REF \
  -V c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x.vcf \
  -o c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.vcf

grep -v '#' c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.vcf | wc -l #2407070749

```

###Get substitutions between ll and lr. Build allele counts file with all positions in ll and lr, and mark all substitutions.

```{r Get substitutions between ll and lr, eval=FALSE, engine='bash'}

#Extract appropriate columns from the VCF:
cd $V_PATH
grep -v '#' c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.vcf | cut -f1,2,4,5,10,11 | awk '{printf ("%s\t%s\t%s\t%s\t%.3s\t%.3s\n", $1, $2, $3, $4, $5, $6)}' > c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions

#Classify the positions into the following categories: NA (whenever there is missing data), multiallelic (whenever there's a comma separating two or more alternative alleles), INDEL (when any allele covers more than one position), invariable (when the reference allele is fixed in both ll and lr), substitution (when alleles are alternatively fixed in ll and lr), substitution with respect to Iberian lynx (when the alternative allele is fixed in both ll and lr), SNP (the remaining cases).
cd $V_PATH
rm c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class
NAMES="Scaffold Position Reference Alternative c_ll_ki_0090 c_lr_zz_0001 Class"
echo $NAMES > c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class
awk '{
if ($5 == "./." || $6 == "./.")
	print $0 "\t NA";
else if ($4 ~ /,/) 
	print $0 "\t Multiallelic";
else if ((length($3) > 1) || (length($4) > 1) || $4 ~ /*/)
	print $0 "\t INDEL";
else if ($5 == "0/0" && $6 == "0/0")
  print $0 "\t Invariable";
else if (($5 == "0/0" && $6 == "1/1") || ($5 == "1/1" && $6 == "0/0"))
	print $0 "\t Substitution";
else if ($5 == "1/1" && $6 == "1/1")
  print $0 "\t IL_Substitution"
else
  print $0 "\t SNP"
}' c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions >> c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class


#Exclude all rows with missing data, INDELs and various alleles.
cd $V_PATH
grep -Ev 'NA|Multiallelic|INDEL' c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class > c_ll_ki_0090_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class_callable_universe

```

##Lynx pardinus - Lynx rufus.
###Obtain joint VCF for pardinus and rufus. Extract data for the two species in order to catalogue substitutions.

```{r Obtain joint VCF for eurasian and rufus, eval=FALSE, engine='bash'}

#Obtain VCF for the eurasian and the rufus:
cd $V_PATH
screen -S c_lp_sm_0298_plus_c_lr_zz_0001_VCF.log
script c_lp_sm_0298_plus_c_lr_zz_0001_VCF.log

$BCF view -s c_lp_sm_0298,c_lr_zz_0001 -Ov -o $V_PATH/c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x.vcf $V_PATH/c_lp_ll_lc_lr_all_sites_renamed.vcf

grep -v '#' c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x.vcf | wc -l #2412763898

```

###Extract SNPs from the VCF. Subset the VCF files in order to keep only SNP variants.

```{r Extract SNPs from the VCF, eval=FALSE, engine='bash'}

#During this step, all INDELs will be dropped from the VCF (not entirely true... INDELs where one variant is monoallelic will remain).
cd $V_PATH
screen -S c_lp_sm_0298_plus_c_lr_zz_0001_VCF_SNPs.log
script c_lp_sm_0298_plus_c_lr_zz_0001_VCF_SNPs.log

java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -Xms16g -Xmx32g -jar $GATK \
  -T SelectVariants \
  --selectTypeToExclude INDEL \
  -R $REF \
  -V c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x.vcf \
  -o c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.vcf

grep -v '#' c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.vcf | wc -l #2407070749

```

###Get substitutions between lp and lr. Build allele counts file with all positions in lp and lr, and mark all substitutions.

```{r Get substitutions between ll and lr, eval=FALSE, engine='bash'}

#Extract appropriate columns from the VCF:
cd $V_PATH
screen -S c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class.log
script c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class.log

grep -v '#' c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.vcf | cut -f1,2,4,5,10,11 | awk '{printf ("%s\t%s\t%s\t%s\t%.3s\t%.3s\n", $1, $2, $3, $4, $5, $6)}' > c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions

#Classify the positions into the following categories: NA (whenever there is missing data), multiallelic (whenever there's a comma separating two or more alternative alleles), INDEL (when any allele covers more than one position), invariable (when the reference allele is fixed in both lp and lr), substitution (when alleles are alternatively fixed in lp and lr), substitution with respect to Iberian lynx (when the alternative allele is fixed in both ll and lr), SNP (the remaining cases).
cd $V_PATH
rm c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class
NAMES="Scaffold Position Reference Alternative c_lp_sm_0298 c_lr_zz_0001 Class"
echo $NAMES > c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class
awk '{
if ($5 == "./." || $6 == "./.")
	print $0 "\t NA";
else if ($4 ~ /,/) 
	print $0 "\t Multiallelic";
else if ((length($3) > 1) || (length($4) > 1) || $4 ~ /*/)
	print $0 "\t INDEL";
else if (($5 == "0/0" && $6 == "0/0") || ($5 == "1/1" && $6 == "1/1"))
  print $0 "\t Invariable";
else if (($5 == "0/0" && $6 == "1/1") || ($5 == "1/1" && $6 == "0/0"))
	print $0 "\t Substitution";
else
  print $0 "\t SNP"
}' c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions >> c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class


#Exclude all rows with missing data, INDELs and various alleles.
cd $V_PATH
grep -Ev 'NA|Multiallelic|INDEL' c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class > c_lp_sm_0298_plus_c_lr_zz_0001_recal_round-1_25x_SNPs.positions_class_callable_universe

```


##Lynx lynx - Lynx pardinus.
###Obtain joint VCF for eurasian and pardinus. Extract data for the two species in order to catalogue substitutions.

```{r Obtain joint VCF for eurasian and rufus, eval=FALSE, engine='bash'}

#Obtain VCF for the eurasian and pardinus:
cd $V_PATH
$BCF view -s c_ll_ki_0090,c_lp_sm_0298 -Ov -o $V_PATH/c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x.vcf $V_PATH/c_lp_ll_lc_lr_all_sites_renamed.vcf

grep -v '#' c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x.vcf | wc -l #2412763898

```

###Extract SNPs from the VCF. Subset the VCF files in order to keep only SNP variants.

```{r Extract SNPs from the VCF, eval=FALSE, engine='bash'}

#During this step, all INDELs will be dropped from the VCF (not entirely true... INDELs where one variant is monoallelic will remain).
cd $V_PATH
java -XX:MaxMetaspaceSize=1g -XX:+UseG1GC -XX:+UseStringDeduplication -Xms16g -Xmx32g -jar $GATK \
  -T SelectVariants \
  --selectTypeToExclude INDEL \
  -R $REF \
  -V c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x.vcf \
  -o c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.vcf

grep -v '#' c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.vcf | wc -l #2407070749

```

###Get substitutions between ll and lr. Build allele counts file with all positions in ll and lr, and mark all substitutions.

```{r Get substitutions between ll and lr, eval=FALSE, engine='bash'}

#Extract appropriate columns from the VCF:
cd $V_PATH
grep -v '#' c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.vcf | cut -f1,2,4,5,10,11 | awk '{printf ("%s\t%s\t%s\t%s\t%.3s\t%.3s\n", $1, $2, $3, $4, $5, $6)}' > c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions

#Classify the positions into the following categories: NA (whenever there is missing data), multiallelic (whenever there's a comma separating two or more alternative alleles), INDEL (when any allele covers more than one position), invariable (when the reference allele is fixed in both ll and lr), substitution (when alleles are alternatively fixed in ll and lr), common substitution with respect to the ancestor (when the alternative allele is fixed in both ll and lr), SNP (the remaining cases).
cd $V_PATH
rm c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions_class
NAMES="Scaffold Position Reference Alternative c_ll_ki_0090 c_lp_sm_0298 Class"
echo $NAMES > c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions_class
awk '{
if ($5 == "./." || $6 == "./.")
	print $0 "\t NA";
else if ($4 ~ /,/) 
	print $0 "\t Multiallelic";
else if ((length($3) > 1) || (length($4) > 1) || $4 ~ /*/)
	print $0 "\t INDEL";
else if ($5 == "0/0" && $6 == "0/0")
  print $0 "\t Invariable";
else if (($5 == "0/0" && $6 == "1/1") || ($5 == "1/1" && $6 == "0/0"))
	print $0 "\t Substitution";
else if ($5 == "1/1" && $6 == "1/1")
  print $0 "\t Comm_substitution"
else
  print $0 "\t SNP"
}' c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions >> c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions_class


#Exclude all rows with missing data, INDELs and various alleles.
cd $V_PATH
grep -Ev 'NA|Multiallelic|INDEL' c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions_class > c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions_class_callable_universe

grep 'Substitution' c_ll_ki_0090_plus_c_lp_sm_0298_recal_round-1_25x_SNPs.positions_class_callable_universe | wc -l #4410056


```
